Limited physiological plasticity and reduced relative variance in physiological rates of ectotherm populations under climate change

### 2024-07-08 ms.qmd
UPdated and re-rendered MS

^1^ Division of Ecology and Evolution, Research School of Biology, The
Australian National University, Canberra, ACT 2600, Australia
\
^2^ Ecology and Evolution Research Centre, School of Biological, Earth
and Environmental Sciences, University of New South Wales, Sydney, NSW,
Australia
\
^3^ School of Life and Environmental Sciences, University of Sydney,
Sydney, NSW 2006, Australia
\
^4^ Department of Biology, Lancaster University, Liverpool, UK
\
$\dagger$ contributed equally
\
$\ddagger$ corresponding author, daniel.noble\@anu.edu.au

@@

Changes in variability of traits governing energy demand are expected to
have important consequences on the flow of energy within and between
populations, communities, and ecosystems [@hendry2016key;
@seebacher2023can; @barneche2021warming; @Sanderson2023; @Bolnick2011].
More variable populations, genetically and/or phenotypically, are
predicted to be associated with broader niches, reduced intraspecific
competition, increased growth rate, decreased vulnerability to
environmental change and lower extinction risk [@forsman2015rethinking;
@forsman2014effects; @Bolnick2011; @hendry2016key; @hart2016variation].
In addition, if phenotypic and genetic variation are correlated and
traits are linked to fitness, changes in variation can limit responses
to selection impacting a populations capacity to evolve
[@Pelletier2012]. Therefore, maintaining intrapopulation variability in
physiological rates in a warmer world may be important for population
resilience to climate change.

@@

```{r, fig-concept}
#| label: fig-concept
#| fig-width: 8
#| fig-height: 12
#| fig-cap: "Calculations of acute and acclimation effect sizes. (A) Two idealised thermal performance curves for animals held at 'cold' ('blue') temperatures and warm ('red') temperatures. Shaded blue and red areas are the thermal optima of the performance curves. Physiological rates are measured for a sample of ectotherms at two different temperatures along the thermal performance curves ($T_{1}$ = 20$^\\circ$C and $T_{2} = 30$^\\circ$C) for both curves. At each temperature a mean physiological rate (R) (points) and its standard deviation (SD) (error bars above and below mean) are estimated. R1.1 and R1.2 are the rates and associated SD (subscripted) for the cold acclimated animals at temperature 1 and 2, respectively. R2.1 and R2.2 are the rates and associated SD (subscripted) for the warm acclimated animals at temperature 1 and 2, respectively. An example of how acute and acclimation $lnCVR_{Q_{10}}$ are calculated from the treatments within the study is provided on the right-hand side of the figure with reference to each of the four possible groups. Note that, two acute effect sizes can be calculated, one for the cold acclimated animals and one for the warm acclimated animals. (B) Species are expected, *a priori*, to vary in their thermal performance curves (thin lines) around an average (thick black line). We restricted our data to areas of each species' performance curve that fell within the natural thermal range of the species (thick lines on each species-level curve). However, given it was not possible to measure the full performance curve for each species some test temperatures within studies may have converged on or moved past the thermal optima. In such cases, we expected our $Q_{10}$ effect sizes to be smaller as indicated by comparing the black dashed lines to grey dashed lines."

fig_concept <- magick::image_read("output/figures/fig1_concept_full.png")
plot(fig_concept)
```

@@

Where, $R_{1}$ and $R_{2}$ are mean physiological rates at temperatures
$T_{1}$ and $T_{2}$, respectively. Log transformation of this ratio
makes the effect size normally distributed. @eq-lnq10 is essentially a
temperature corrected equivalent to the log response ratio (lnRR)
[@Hedges1999-de; @Lajeunesse2011-fq] when the numerator and denominator
are measured at different temperatures. This allows comparisons of the
means from two temperature treatments directly regardless of the
absolute measurement temperatures. The sampling variance for @eq-lnq10
can be computed as follows (as described in @Noble2022-ty):

@@

Here, $SD_{1}^2$ and $SD_{2}^2$ are the standard deviations, and $N_{1}$
and $N_{2}$ are the sample sizes, of the groups measured at $T_{1}$ and
$T_{2}$, respectively (@fig-concept A).

@@

$R_{1}$, $R_{2}$, $SD_{1}^2$ and $SD_{2}^2$ can all be taken from
samples of organisms measured acutely at two temperatures or after
having been acclimated these same temperatures (@fig-concept A). For
studies that measure acute and acclimated responses we used the mean,
standard deviation, and sample size to derive both acute and acclimation
$lnRR_{Q_{10}}$, $lnVR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ estimates. For
studies that only measured $R_{1}$, $R_{2}$, $SD_{1}^2$ and $SD_{2}^2$
acutely we could only calculate acute versions of these effect size
estimates. For all effect sizes the higher temperature was in the
numerator and the lower of the two temperatures in the denominator. As
such, positive effect sizes suggest that the mean (i.e.,
$lnRR_{Q_{10}}$) or relative variance (i.e., $lnCVR_{Q_{10}}$) is larger
at the higher of the two temperatures (numerator) when standardized to
10°C. Importantly, $lnRR_{Q_{10}}$, $lnCVR_{Q_{10}}$ and
$lnVR_{Q_{10}}$, as with $Q_{10}$ more generally, all assume that the
effect of temperature on physiological rates (or changes in relative
variance) is log-linear (see *Supplementary Materials* for further
discussion). We test and control for any violations of these assumptions
in our analysis (see below) ([@fig-concept]B).

@@

# Load and plot Figure 1. Note that this figure was derived using 3 separate code chunks described below. Figures were exported, imported into Adobe Illustrator for minor formatting changes and grid organisation. 
  fig2 <- image_read("./output/figures/Fig2_r1.png")
  plot(fig2)

@@

Across habitats the extent to which whole-organism versus biochemical
processes acclimated varied ("Model 3"; [@fig-btrait]A-C; and see
Supplement, @fig-S2) for more detailed trait analysis). Biochemical
processes were more effective at acclimating compared to whole-organism
traits for ectotherms occupying marine habitats ([@fig-btrait]B),
whereas both whole-organism and biochemical traits acclimated in
freshwater ectotherms ([@fig-btrait]C). Unsurprisingly, neither broad
trait category showed evidence of acclimation in terrestrial ectotherms
([@fig-btrait]A).

@@

```{r, fig-2}
#| label: fig-2
#| fig-cap: "A Mean acute and acclimation $lnRR_{Q_{10}}$ across ectotherms in marine, freshwater, and terrestrial habitats. B Mean acute and acclimation $lnCVR_{Q_{10}}$ across ectotherms in marine, freshwater and terrestrial habitats. For both plots, k = total number of effect size estimates while the numbers in brackets indicate the number of species. Sample sizes are the same for panel A and B. In panels A and B, thick bars are 95% confidence intervals (CI) and thin bars 95% prediction intervals (PI). $\\Beta$ values are the contrasts between acute and acclimation means within each habitat. $\\mu$ values are the overall meta-analytic means averaged across acute and acclimation types within each habitat type. In both cases, their 95% CI's are indicated within square brackets. $p_{MCMC}$ values are the posterior probability of the contrast or overall meta-analytic mean being different from zero. For ease of visualisation, all the raw data plotted for both acute and acclimation type effect sizes are presented as circles."
#| warning: false
#| echo: false

@@

  fig3 <- (p3a + p3b) + plot_annotation(tag_levels = "A")
  ggsave(fig3, filename = "./output/figures/fig3_r1.png", width = 21.88, height = 9)

  fig3 <- magick::image_read("./output/figures/fig3_r1.png")
  fig3
```

@@

Across habitats whole-organism versus biochemical processes showed a
similar decrease in relative variance in physiological rates
([@fig-btrait]D-F; see Supplement @fig-S2) for more detailed trait
analysis), except for biochemical processes in ectotherms occupying
freshwater habitats, which showed small, albeit significant increase in
relative variance when acclimated ([@fig-btrait]F). Reduced relative
variance was particularly prominent for resting metabolic rates and
sprint speed although traits differed in whether they exhibited a
reduction in variation in physiological rates at high temperatures (see
*Supplemental Material*, @fig-S2) ("Model 3"). Each life-history stage
exhibited the same pattern of relative variance change in each of the
habitats ("Model 4"; @fig-cvrlh).

@@

############
# Full fig
############
figbtrait <-(p3 | p1 | p2) / (p3_cvr | p1_cvr | p2_cvr) + plot_annotation(tag_levels = "A") 
ggsave(figbtrait, filename = "./output/figures/fig4_btrait_r1.png", width = 28, height = 14, device = "png")

figbtrait <- magick::image_read("./output/figures/fig4_btrait_r1.png")
figbtrait
```

@@

fig4 <-(p3 | p1 | p2) / (p3_cvr | p1_cvr | p2_cvr) + plot_annotation(tag_levels = "A") 
ggsave(fig4, filename = "./output/figures/fig4_lh_r1.pdf", width = 28, height = 14, device = "pdf")

fig4 <- magick::image_read("./output/figures/fig4_lh_r1.pdf")
fig4
```

@@

  ggsave(p_cv, width = 18, height = 10, filename = "./output/figures/fig_p_cv_r1.png")

  p_cv <- image_read("./output/figures/fig_p_cv_r1.png")
  p_cv 
```

@@

ggsave(p, filename = "./output/figures/fig-preds.png", width = 7.8, height = 10.1)

image_read("./output/figures/fig-preds.png")
```

@@

We show that acclimation of physiological rates and changes in relative
variance differ across habitats. Our findings confirm previous results
that quantify the different capacity of terrestrial, marine and
freshwater ectotherms to acclimate [@Seebacher2015-wi;
@Gunderson2015-hz; @morley2019physiological]. Our analysis generally
supports that of @Seebacher2015-wi, @Gunderson2015-hz and
@morley2019physiological that all show a general inability of
terrestrial ectotherms to physiologically acclimate. The change in
acclimation $Q_{10}$ we found was similar to @Seebacher2015-wi for
freshwater organisms (\~17%), but higher in marine ectotherms (decrease
of 16% versus \~10% in @Seebacher2015-wi), and opposite in terrestrial
ectotherms (increase of \~6% compared to an \~8% decrease in
@Seebacher2015-wi), possibly due to the use of new $Q_{10}$ based effect
sizes that allow us to control for sampling variance. Greater capacity
for acclimation in aquatic organisms may be the result of fewer
opportunities for behavioural thermoregulation in aquatic environments
making physiological remodeling important for maintaining homeostasis
[@Gunderson2015-hz; @morley2019physiological]. While marine and
freshwater ectotherms were capable of partial acclimation it is still
important to recognise that, on average, the effect size was small
(amounting to $Q_{10}$ dropping from \~1.8 to 1.6).

@@

Our results further highlight the potential vulnerability of terrestrial
ectotherms to climate change, while calling into question if acclimation
in marine systems will be able to counteract the potentially negative
effects of reductions in relative variance. Assuming changes in relative
variation in physiological rates are underpinned by genetic variation,
and there is a genetic correlation with fitness, such reductions in
physiological variance could limit a populations ability to adapt to
future climates [@urban2023and]. Of course, considering how behavioural
modifications buffer organisms from high temperatures is also needed.
Terrestrial species likely have greater opportunity for behavioural
microhabitat selection in thermally heterogeneous environments which is
expected to decrease the impacts of climate change [@huey2012predicting;
@sunday2014thermal]. Importantly, while we seen a reduction in relative
variation, absolute variation in physiological rates did increase with
temperature (@fig-s1) -- albeit to a much lesser extent in terrestrial
and marine habitats compared to freshwater -- supporting the general
notion that capacity for adaptation may still be limited in terrestrial
and marine habitats. As of yet, it is unclear whether a focus on
relative or absolute variance should be prioritized. Future theoretical
models are needed to better understand the ecological and evolutionary
consequences of changes in both mean and variance.

@@

<br> 
<div id ="refs"></div>
<br>

## Supplemental Results and Figures

@@

prisma <- magick::image_read("./output/figures/PRISMA.png")
prisma
```

@@

ggsave(lnrrq10_temp, file = "./output/figures/lnrrq10_temp.png", height = 6, width = 8)

lnrrq10_temp <- magick::image_read("./output/figures/lnrrq10_temp.png")
lnrrq10_temp

@@

  ggsave(accl_time, file = "./output/figures/accl_time.png", height = 6, width = 8.5)

accl_time <- magick::image_read("./output/figures/accl_time.png")
accl_time
```

@@

col_acute <- c("#3F5FAC", "#EF3D25")
orchard_plot(model1.s1_int, mod = "name", by = "habitat", xlab = TeX("$lnRR_{Q_{10}}$"), group = "spp", angle = 45, trunk.size = 1.8, branch.size = 2.8, alpha = 0.4, condition.lab = "Habitat") + scale_x_discrete(labels = c("LnRR_Q10_acute_t2" = "Warm - Acute", "LnRR_Q10_acute_t1" = "Cool - Acute")) + scale_fill_manual(values = col_acute) + scale_colour_manual(values = col_acute) + ylim (-0.5, 2.5) +  theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_text(angle = 45, size = 24),
                                       legend.title = element_text(size = 30),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24), legend.position = "top") 
                               #annotate("text", label= paste0("k = ", k_lnrr_acute$k, " (", k_lnrr_acute$sdy, ")"), x = c(1.2, 2.2), y = 2)
```

@@

@fig-lnrrtrait; @fig-cvrtrait

@@

fig2 <- (ptm / ptf / ptt) + plot_annotation(tag_levels = "A")
ggsave(fig2, filename = "./output/figures/fig2_trait.png", height = 20, width = 15)

img_fig2 <- magick::image_read("./output/figures/fig2_trait.png")
img_fig2
```

@@

figcvrtrait <- (ptm / ptf / ptt) + plot_annotation(tag_levels = "A", tag_suffix = ")")
ggsave(figcvrtrait, filename = "./output/figures/figcvrtrait.png", height = 20, width = 15)

img_figcvrtrait <- magick::image_read("./output/figures/figcvrtrait.png")
img_figcvrtrait
```

@@

We explored mean-variance relationships for the acute and acclimation
responses across all traits and habitats. We estimated the scaling
relationship between log standard deviation in physiological rates
\[log(SD)\] and log mean physiological rates \[log(mean)\], accounting
for non-independence resulting from effects coming from the same
species, study and traits (i.e., random effects of species, study and
trait) as done in our main analyses. We also included an interaction
between log(mean) and habitat type to better understand how the scaling
relationship between log(SD) and log(mean) varies across habitats.

Overall, we found that the relationship between log(mean) and log(SD) of
the acute and acclimation responses was generally linear, suggesting
that the use of $lnCVR_{Q_{10}}$ was justified (@fig-mv). Overall, the
scaling relationship between log(SD) and log(mean) was sub-linear across
all habitats (@tbl-slopes), however, ectotherms from terrestrial
habitats had much shallower slopes than marine and freshwater
ectotherms, particularly at higher treatment temperatures, indicating
increased mean physiological rates generally do not result in higher
between individual variance in physiological rates (@tbl-slopes).
Interestingly, in marine ectotherms the slope was highest at cooler
temperatures, wheeras the slope was suppressed when acclimated and/or
measured at higher temperatures (i.e., r.1. compared to r1.2, r2.1,
r2.2) (@tbl-slopes). In freshwater ectotherms, there we some differences
in scaling relationships but they were all fairly comparable no matter
what acclimatin and test temperature (@tbl-slopes).

@@

 p_mv <-  (mv1 / mv2)  + plot_annotation(tag_levels = "A") + theme(plot.tag = element_text(size = 20))
  ggsave(p_mv, file = "./output/figures/plot_mv.png", height = 8, width = 8)

  p_mv <- magick::image_read("./output/figures/plot_mv.png")
  p_mv
```

@@

  ggsave(p3, filename = "./output/figures/figS1.pdf", width = 11.5, height = 8)

  figs1 <- image_read("./output/figures/figS1.pdf")
  figs1
```

@@

figs3 <- (ps3_1 | ps3_2 | ps3_3) + plot_annotation(tag_levels = "A", tag_suffix = ")") & theme(plot.tag = element_text(size = 20))
ggsave(figs3, width = 16, height = 5.7, filename = "./output/figures/figs3.png")

figs3 <- image_read("./output/figures/figs3.png")
figs3
```

@@

ggsave(p_acf, width = 18, height = 10, filename = "./output/figures/fig_p_acf_r1.pdf")

  p_acf <- image_read("./output/figures/fig_p_acf_r1.pdf")
  p_acf 
```

@@

figs2 <- (ps2_1 | ps2_2 | ps2_3) + plot_annotation(tag_levels = "A", tag_suffix = ")") & theme(plot.tag = element_text(size = 20))
ggsave(figs2, width = 11.5, height = 4.8, filename = "./output/figures/figS2.png")

figs2 <- image_read("./output/figures/figS2.png")
figs2
```

>>>>

^1^ Division of Ecology and Evolution, Research School of Biology, The
Australian National University, Canberra, ACT 2600, Australia\
^2^ Ecology and Evolution Research Centre, School of Biological, Earth
and Environmental Sciences, University of New South Wales, Sydney, NSW,
Australia\
^3^ School of Life and Environmental Sciences, University of Sydney,
Sydney, NSW 2006, Australia\
^4^ Department of Biology, Lancaster University, Liverpool, UK\
$\dagger$ contributed equally\
$\ddagger$ corresponding author, daniel.noble\@anu.edu.au

@@

Changes in variability of traits governing energy demand are expected to
have important consequences on the flow of energy within and between
populations, communities, and ecosystems [@hendry2016key;
@seebacher2023can; @barneche2021warming; @Sanderson2023; @Bolnick2011].
More variable populations, genetically and/or phenotypically, are also 
predicted to be associated with broader niches, have increased growth rate, decreased vulnerability to
environmental change and lower extinction risk [@forsman2015rethinking;
@forsman2014effects; @Bolnick2011; @hendry2016key; @hart2016variation].
In addition, if phenotypic and genetic variation are correlated and
traits are linked to fitness, changes in variation may limit responses
to selection impacting a populations capacity to evolve
[@Pelletier2012]. Therefore, maintaining intrapopulation variability in
physiological rates in a warmer world may be important for population
resilience to climate change.

@@

```{r, fig-concept}
#| label: fig-concept
#| fig-width: 8
#| fig-height: 12
#| fig-cap: "Calculations of acute and acclimation effect sizes. (A) Two idealised thermal performance curves for animals held at 'cold' ('blue') temperatures and warm ('red') temperatures. Shaded blue and red areas are the thermal optima of the performance curves. Physiological rates are measured for a sample of ectotherms at two different temperatures along the thermal performance curves ($T_{1}$ = 20°C and $T_{2}$ = 30°C) for both curves. At each temperature a mean physiological rate (R) (points) and its standard deviation (SD) (error bars above and below mean) are estimated. R1.1 and R1.2 are the rates and associated SD (subscripted) for the cold acclimated animals at temperature 1 and 2, respectively. R2.1 and R2.2 are the rates and associated SD (subscripted) for the warm acclimated animals at temperature 1 and 2, respectively. An example of how acute and acclimation $lnCVR_{Q_{10}}$ are calculated from the treatments within the study is provided on the right-hand side of the figure with reference to each of the four possible groups. Note that, two acute effect sizes can be calculated, one for the cold acclimated animals and one for the warm acclimated animals. (B) Species are expected, *a priori*, to vary in their thermal performance curves (thin lines) around an average (thick black line). We restricted our data to areas of each species' performance curve that fell within the natural thermal range of the species (thick lines on each species-level curve). However, given it was not possible to measure the full performance curve for each species some test temperatures within studies may have converged on or moved past the thermal optima. In such cases, we expected our $Q_{10}$ effect sizes to be smaller as indicated by comparing the black dashed lines to grey dashed lines."

fig_concept <- magick::image_read("output/figures/fig1.png")
plot(fig_concept)
```

@@

Where, $R_{1}$ and $R_{2}$ are mean physiological rates at temperatures
$T_{1}$ and $T_{2}$, respectively. Log transformation of this ratio
makes the effect size normally distributed. @eq-lnq10 is essentially a
temperature corrected equivalent to the log response ratio (lnRR)
[@Hedges1999-de; @Lajeunesse2011-fq] when the numerator and denominator
are measured at different temperatures. This allows comparisons of the
means from two temperature treatments directly regardless of the
absolute measurement temperatures. The sampling variance for @eq-lnq10
can be computed as follows [as described in @Noble2022-ty]:

@@

Here, $SD_{1}^2$ and $SD_{2}^2$ are the standard deviations, and $N_{1}$
and $N_{2}$ are the sample sizes, of the groups measured at $T_{1}$ and
$T_{2}$, respectively ([@fig-concept]A).

@@

$R_{1}$, $R_{2}$, $SD_{1}^2$ and $SD_{2}^2$ can all be taken from
samples of organisms measured acutely at two temperatures or after
having been acclimated these same temperatures ([@fig-concept]A). For
studies that measure acute and acclimated responses we used the mean,
standard deviation, and sample size to derive both acute and acclimation
$lnRR_{Q_{10}}$, $lnVR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ estimates. For
studies that only measured $R_{1}$, $R_{2}$, $SD_{1}^2$ and $SD_{2}^2$
acutely we could only calculate acute versions of these effect size
estimates. For all effect sizes the higher temperature was in the
numerator and the lower of the two temperatures in the denominator. As
such, positive effect sizes suggest that the mean (i.e.,
$lnRR_{Q_{10}}$) or relative variance (i.e., $lnCVR_{Q_{10}}$) is larger
at the higher of the two temperatures (numerator) when standardized to
10°C. Importantly, $lnRR_{Q_{10}}$, $lnCVR_{Q_{10}}$ and
$lnVR_{Q_{10}}$, as with $Q_{10}$ more generally, all assume that the
effect of temperature on physiological rates (or changes in relative
variance) is log-linear (see [@fig-concept]B & *Supplementary Materials* for further
discussion). We test and control for any violations of these assumptions
in our analysis (see below).

@@

# Load and plot Figure 1. Note that this figure was derived using 3 separate code chunks described below. Figures were exported, imported into Adobe Illustrator for minor formatting changes and grid organisation. 
  fig2 <- image_read("./output/figures/Fig2.png")
  plot(fig2)

@@

Across habitats the extent to which whole-organism versus biochemical
processes acclimated varied ("Model 3"; [@fig-btrait]A-C; and see
*Supplementary Materials*, @fig-lnrrtrait for more detailed trait analysis). Biochemical
processes were more effective at acclimating compared to whole-organism
traits for ectotherms occupying marine habitats ([@fig-btrait]B),
whereas both whole-organism and biochemical traits acclimated in
freshwater ectotherms ([@fig-btrait]C). Unsurprisingly, neither broad
trait category showed evidence of acclimation in terrestrial ectotherms
([@fig-btrait]A).

@@

```{r, fig-2}
#| label: fig-2
#| fig-cap: "A) Mean acute and acclimation $lnRR_{Q_{10}}$ across ectotherms in marine, freshwater, and terrestrial habitats. B) Mean acute and acclimation $lnCVR_{Q_{10}}$ across ectotherms in marine, freshwater and terrestrial habitats. For both plots, k = total number of effect size estimates while the numbers in brackets indicate the number of species. Sample sizes are the same for panel A) and B). In panels A) and B), thick bars are 95% confidence intervals (CI) and thin bars 95% prediction intervals (PI). $\\Beta$ values are the contrasts between acute and acclimation means within each habitat. $\\mu$ values are the overall meta-analytic means averaged across acute and acclimation types within each habitat type. In both cases, their 95% CI's are indicated within square brackets. $p_{MCMC}$ values are the posterior probability of the contrast or overall meta-analytic mean being different from zero. For ease of visualisation, all the raw data plotted for both acute and acclimation type effect sizes are presented as circles."
#| warning: false
#| echo: false

@@

  fig3 <- (p3a + p3b) + plot_annotation(tag_levels = "A")
  ggsave(fig3, filename = "./output/figures/fig3.png", width = 21.88, height = 9)

  fig3 <- magick::image_read("./output/figures/fig3.png")
  fig3
```

@@

Across habitats whole-organism versus biochemical processes showed a
similar decrease in relative variance in physiological rates
([@fig-btrait]D-F; see *Supplementary Materials*, @fig-cvrtrait) for more detailed trait
analysis), except for biochemical processes in ectotherms occupying
freshwater habitats, which showed small, albeit significant increase in
relative variance when acclimated ([@fig-btrait]F). Reduced relative
variance was particularly prominent for resting metabolic rates and
sprint speed although traits differed in whether they exhibited a
reduction in variation in physiological rates at high temperatures (see
*Supplemental Material*, @fig-cvrtrait) ("Model 3"). Each life-history stage
exhibited the same pattern of relative variance change in each of the
habitats ("Model 4"; @fig-cvrlh).

@@

############
# Full fig
############
figbtrait <-(p3 | p1 | p2) / (p3_cvr | p1_cvr | p2_cvr) + plot_annotation(tag_levels = "A") 
ggsave(figbtrait, filename = "./output/figures/fig4.png", width = 28, height = 14, device = "png")

figbtrait <- magick::image_read("./output/figures/fig4.png")
figbtrait
```

@@

fig4 <-(p3 | p1 | p2) / (p3_cvr | p1_cvr | p2_cvr) + plot_annotation(tag_levels = "A") 
ggsave(fig4, filename = "./output/figures/fig5.pdf", width = 28, height = 14, device = "pdf")

fig4 <- magick::image_read("./output/figures/fig5.pdf")
fig4
```

@@

  ggsave(p_cv, width = 18, height = 10, filename = "./output/figures/fig6.png")

  p_cv <- image_read("./output/figures/fig6.png")
  p_cv 
```

@@

ggsave(p, filename = "./output/figures/fig7.png", width = 7.8, height = 10.1)

image_read("./output/figures/fig7.png")
```

@@

We show that acclimation of physiological rates and changes in relative
variance differ across habitats. Our findings confirm previous results
that quantify the different capacity of terrestrial, marine and
freshwater ectotherms to acclimate [@Seebacher2015-wi;
@Gunderson2015-hz; @morley2019physiological]. Our analysis generally
supports that of @Seebacher2015-wi, @Gunderson2015-hz and
@morley2019physiological that all show a general inability of
terrestrial ectotherms to physiologically acclimate. The change in
acclimation $Q_{10}$ we found was similar to @Seebacher2015-wi for
freshwater organisms (~17%), but higher in marine ectotherms [decrease
of 16% versus ~10% in @Seebacher2015-wi], and opposite in terrestrial
ectotherms [increase of ~6% compared to an ~8% decrease in
@Seebacher2015-wi], possibly due to the use of newly derived
$Q_{10}$ effect sizes that allow us to control for sampling variance.
Greater capacity for acclimation in aquatic organisms may be the result
of fewer opportunities for behavioural thermoregulation in aquatic
environments making physiological remodeling important for maintaining
homeostasis [@Gunderson2015-hz; @morley2019physiological].
Additionally, aquatic and terrestrial environments can vary in
nutritional resources [e.g., fatty acid content of prey;
@twining2019aquatic] that can impact physiological rates and their
variability by affecting performance curves and acclimation responses
[@hardison2024diet]. While marine and freshwater ectotherms were
capable of partial acclimation it is still important to recognise that,
on average, the effect size was small (amounting to $Q_{10}$ dropping
from ~1.8 to 1.6).

@@

Our results further highlight the vulnerability of terrestrial
ectotherms to climate change, while calling into question if acclimation
in marine systems will be able to counteract the potentially negative
effects of reductions in relative variance. Assuming changes in relative
variation in physiological rates are underpinned by genetic variation,
and there is a genetic correlation with fitness, such reductions in
physiological variance could limit a populations ability to adapt to
future climates [@urban2023and]. Of course, considering how behavioural
modifications buffer organisms from high temperatures is also needed.
Terrestrial species likely have greater opportunity for behavioural
microhabitat selection in thermally heterogeneous environments which is
expected to decrease the impacts of climate change [@huey2012predicting;
@sunday2014thermal]. Importantly, while we seen a reduction in relative
variation, absolute variation in physiological rates did increase with
temperature (@fig-s1) -- albeit to a much lesser extent in terrestrial
and marine habitats compared to freshwater -- supporting the general
notion that capacity for adaptation may still be limited in terrestrial
and marine habitats. As of yet, it is unclear whether a focus on
relative or absolute variance should be prioritized. Future theoretical
models are needed to better understand the ecological and evolutionary
consequences of changes in both mean and variance.

@@

<br>

::: {#refs}
:::

<br>

## Supplemental Materials

@@

prisma <- magick::image_read("./output/figures/figS1.png")
prisma
```

@@

ggsave(lnrrq10_temp, file = "./output/figures/figS2.png", height = 6, width = 8)

lnrrq10_temp <- magick::image_read("./output/figures/figS2.png")
lnrrq10_temp

@@

  ggsave(accl_time, file = "./output/figures/figS3.png", height = 6, width = 8.5)

accl_time <- magick::image_read("./output/figures/figS3.png")
accl_time
```

@@

col_acute <- c("#3F5FAC", "#EF3D25")
Sacute <- orchard_plot(model1.s1_int, mod = "name", by = "habitat", xlab = TeX("$lnRR_{Q_{10}}$"), group = "spp", angle = 45, trunk.size = 1.8, branch.size = 2.8, alpha = 0.4, condition.lab = "Habitat") + scale_x_discrete(labels = c("LnRR_Q10_acute_t2" = "Warm - Acute", "LnRR_Q10_acute_t1" = "Cool - Acute")) + scale_fill_manual(values = col_acute) + scale_colour_manual(values = col_acute) + ylim (-0.5, 2.5) +  theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_text(angle = 45, size = 24),
                                       legend.title = element_text(size = 30),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24), legend.position = "top") 
                               #annotate("text", label= paste0("k = ", k_lnrr_acute$k, " (", k_lnrr_acute$sdy, ")"), x = c(1.2, 2.2), y = 2)

Sacute
ggsave(Sacute, file = "./output/figures/figS4.png", height = 10, width = 12)
```

@@

Acclimation capacity varied across trait categories and habiats with measures of resting metabolic rate, including associated biochemical reactions like oxidative phosphorylation (OXPHOS) and ATPase activity, acclimating in marine and freshwater ectotherms (@fig-lnrrtrait). Whether relative variation in physiological rates changes also depended on trait type, with freshwater ectotherms generally maintaining relative variance in physiological rates better than marine and freshwater ectotherms (@fig-cvrtrait). We note though that some traits have very small sample sizes on their own and should be interpreted with caution.

@@

fig2 <- (ptm / ptf / ptt) + plot_annotation(tag_levels = "A")
ggsave(fig2, filename = "./output/figures/figS5.png", height = 20, width = 15)

img_fig2 <- magick::image_read("./output/figures/figS5.png")
img_fig2
```

@@

figcvrtrait <- (ptm / ptf / ptt) + plot_annotation(tag_levels = "A", tag_suffix = ")")
ggsave(figcvrtrait, filename = "./output/figures/figS6.png", height = 20, width = 15)

img_figcvrtrait <- magick::image_read("./output/figures/figS6.png")
img_figcvrtrait
```

@@

We explored mean-variance relationships for the acute and acclimation
responses across all traits and habitats. We estimated the scaling
relationship between log standard deviation in physiological rates
[log(SD)] and log mean physiological rates [log(mean)], accounting
for non-independence resulting from effects coming from the same
species, study and traits (i.e., random effects of species, study and
trait) as done in our main analyses. We also included an interaction
between log(mean) and habitat type to better understand how the scaling
relationship between log(SD) and log(mean) varies across habitats.

Overall, we found that the relationship between log(mean) and log(SD) of
the acute and acclimation responses was generally linear, suggesting
that the use of $lnCVR_{Q_{10}}$ was justified (@fig-mv). Overall, the
scaling relationship between log(SD) and log(mean) was sub-linear across
all habitats (@tbl-slopes), however, ectotherms from terrestrial
habitats had much shallower slopes than marine and freshwater
ectotherms, particularly at higher treatment temperatures, indicating
increased mean physiological rates generally do not result in higher
between individual variance in physiological rates (@tbl-slopes).
Interestingly, in marine ectotherms the slope was highest at cooler
temperatures, wheeras the slope was suppressed when acclimated and/or
measured at higher temperatures (i.e., r.1. compared to r1.2, r2.1,
r2.2) (@tbl-slopes). In freshwater ectotherms, there we some differences
in scaling relationships but they were all fairly comparable no matter
what acclimation and test temperature (@tbl-slopes).

@@

 p_mv <-  (mv1 / mv2)  + plot_annotation(tag_levels = "A") + theme(plot.tag = element_text(size = 20))
  ggsave(p_mv, file = "./output/figures/figS7.png", height = 8, width = 8)

  p_mv <- magick::image_read("./output/figures/figS7.png")
  p_mv
```

@@

  ggsave(p3, filename = "./output/figures/figS8.pdf", width = 11.5, height = 8)

  figs1 <- image_read("./output/figures/figS8.pdf")
  figs1
```

@@

figs3 <- (ps3_1 | ps3_2 | ps3_3) + plot_annotation(tag_levels = "A", tag_suffix = ")") & theme(plot.tag = element_text(size = 20))
ggsave(figs3, width = 16, height = 5.7, filename = "./output/figures/figS9.png")

figs3 <- image_read("./output/figures/figS9.png")
figs3
```

@@

ggsave(p_acf, width = 18, height = 10, filename = "./output/figures/figS10.pdf")

  p_acf <- image_read("./output/figures/figS10.pdf")
  p_acf 
```

@@

figs2 <- (ps2_1 | ps2_2 | ps2_3) + plot_annotation(tag_levels = "A", tag_suffix = ")") & theme(plot.tag = element_text(size = 20))
ggsave(figs2, width = 11.5, height = 4.8, filename = "./output/figures/figS11.png")

figs2 <- image_read("./output/figures/figS11.png")
figs2
```
